---
title: "Activity - Reproducible research Assignment 1"
author: "Magnus Ahlström"
date: "14. okt. 2015"
output: html_document
---

##Loading and preprocessing the data

```{r}
setwd("/Users/magnusglindvad/Documents/Rprogramming/RepResearch")
activity = read.csv("RepData_PeerAssessment1/activity.csv")
```

##What is mean total number of steps taken per day?

1) Calculating total number of steps pr. day and plotting it in a histogram
```{r}
tot_steps_day = with(activity, tapply(steps, date, sum, na.rm = T))
hist(tot_steps_day)
```

2) Calcluting mean and median total number of steps pr. day
```{r}
mean = round(mean(tot_steps_day))
median = median(tot_steps_day)
```

- The mean total number of steps pr. day is: `r  mean`
- The median total number of steps pr. day is: `r  median`

##What is the average daily activity pattern?

1)
Calculating and plotting mean steps pr. interval averaged over all days
```{r}
mean_steps_int = with(activity, tapply(steps, interval, mean, na.rm = T))
plot(x = as.numeric(names(mean_steps_int)), 
     y = mean_steps_int,
     type = "l")
```

2)
Calculating the interval with the maximum number of steps
```{r}
max_steps = max(mean_steps_int)
int_max_steps = names(which(mean_steps_int == max_steps))
to_print = paste0(int_max_steps, " - ", (as.numeric(int_max_steps) + 5))
```

- The 5 minute interval of the day that has the maximum average number of steps is: `r to_print`

##Imputing missing values

1) Calculating number of incomplete cases
```{r}
inc.cases = sum(!complete.cases(activity))
```

- The number of incomplete cases in the activity dataset is `r inc.cases`

2) My strategy will be to fill NAs with the average of that interval from all other days as calculated in "What is the average daily activity pattern? - question 1"

3) Creating a function to calculate the missing values, and using it to return a column with the filled in missing values for the activity dataset
```{r}
act.fill.nas = function(x) {
        y = vector()
        for (i in seq_along(x))
                if(is.na(x[i])) {
                      int.val = as.character(activity[,"interval"][i])
                      index = which(names(mean_steps_int) == int.val)
                      y[i] = mean_steps_int[index]
                } else {
                        y[i] = x[i]
                }
        x = y
        x
}
activity$steps = act.fill.nas(activity$steps)
head(activity)
```

4) 
```{r}
tot_steps_day = with(activity, tapply(steps, date, sum, na.rm = T))
hist(tot_steps_day)
mean = format(round(mean(tot_steps_day)), scientific = F)
median = format(round(median(tot_steps_day)), scientific = F)
```

- Now the mean total number of steps pr. day is: `r  mean`
- And the median total number of steps pr. day is: `r  median`
- The Values differ, the median is very close, to the same, but the mean value, changed a lot. In this setting when calculating total steps and ignoring NAs all NAs will be counted as zero. so the results will be biased towards zero total steps pr. day.

```{r}
require(dplyr)
require(ggplot2)
activity = mutate(activity,
                  date = as.Date(date, format = "%Y-%m-%d"),
                  daytype = ifelse(weekdays(date) %in% c("Lørdag", "Søndag"), "weekend", "weekday")) %>%
        group_by(daytype, interval) %>%
        summarise(av_steps = mean(steps))

ggplot(activity, aes(x = interval, y = av_steps)) +
        geom_line() + 
        facet_grid(daytype ~ .)
```

